%%
% BIThesis 研究生学位论文模板 The BIThesis Template for Graduate Thesis
% This file has no copyright assigned and is placed in the Public Domain.
%%

\begin{abstract}
  为满足海洋信息网络和空天地海一体化体系中日益增长的水下信息传输需求，水下无线光通信因其高带宽、低时延等优势，已成为近距离高速水下链路的重要研究方向。然而，不同水下传播场景下的主导失真机制差异显著：在低动态浑浊环境中，多重散射和非视距传播容易引起时延扩展与频率选择性衰落；在高动态移动环境中，平台相对运动又会进一步带来多普勒扩展以及相干接收条件下的相位噪声问题。针对传统 OFDM 体制在上述场景中适应性不足的问题，本文围绕水下无线光通信中的仿射频分复用技术开展研究，重点面向低动态浑浊多径场景和高动态移动场景设计两类具有针对性的传输方案。

  (1)针对低动态水下场景中由多重散射引起的时延扩展与频率选择性衰落问题，本文提出一种面向强度调制/直接检测架构的非对称限幅光仿射频分复用传输方案。该方案将 ACO 机制与 AFDM 波形相结合，通过奇数子载波映射、厄米特对称构造与非对称限幅，在满足非负实信号约束的基础上增强系统对多径传播的适应能力；同时，分析了余弦型 DAFT 核下的参数设计条件，给出了兼顾逆变换稳定性与多径分离能力的啁啾参数设计准则。仿真结果表明，所提方案在低动态多径条件下相较传统光正交频分复用方案具有更好的误码率性能和多径鲁棒性。

  (2)针对高动态水下场景中相对运动引起的多普勒扩展以及相干架构下的相位噪声影响，本文提出一种面向相干接收的 AFDM 索引调制传输方案。该方案建立了包含多径、多普勒频移和维纳相位噪声的系统模型，推导了离散仿射傅里叶变换域的等效输入输出关系，并进一步利用 AFDM-IM 未激活位置所形成的静默结构作为隐式参考，构造基于泄露能量最小化与稀疏度量驱动的盲搜方法，在无额外导频开销下精准补偿主导的残余载波频偏，同时对维纳相位噪声实现鲁棒容忍。仿真结果表明，所提方案在高动态条件下较基准 OFDM-IM 方案具有更优的误码率性能，并在一定激光器线宽范围内表现出较好的相位噪声鲁棒性。

  综上，本文围绕低动态与高动态两类典型水下场景，研究了 AFDM 在水下无线光通信中的波形构造、参数设计与检测方法，说明了 AFDM 面向不同动态条件进行场景化适配的可行性，可为高可靠水下光通信系统设计提供理论参考。  
\end{abstract}

% 如需手动控制换行连字符位置，可写 aa\-bb，详见
% https://bithesis.bitnp.net/faq/hyphen.html

\begin{abstractEn}
  To meet the demand for underwater information transmission in marine information networks and space-air-ground-sea integrated systems, underwater wireless optical communication (UWOC) has become an important solution for short-range and high-rate underwater links due to its high bandwidth and low latency. However, the dominant impairment mechanisms vary significantly across different underwater propagation scenarios. In low-dynamic turbid environments, multiple scattering and non-line-of-sight propagation mainly cause delay spread and frequency-selective fading. In high-dynamic mobile environments, platform motion further introduces Doppler spread and phase noise under coherent reception. To address the limited adaptability of conventional OFDM in these scenarios, this thesis investigates affine frequency division multiplexing (AFDM) for UWOC and develops two dedicated transmission schemes for representative low-dynamic turbid multipath scenarios and high-dynamic mobile scenarios.

  (1)For low-dynamic underwater scenarios affected by delay spread and frequency-selective fading caused by multiple scattering, an asymmetrically clipped optical affine frequency division multiplexing (ACO-AFDM) scheme is proposed for intensity modulation/direct detection (IM/DD) systems. By combining the ACO mechanism with the AFDM waveform through odd-subcarrier mapping, Hermitian symmetry, and asymmetric clipping, the proposed scheme improves multipath adaptability while satisfying the nonnegative real-signal constraint. In addition, the parameter design conditions under the cosine-type DAFT kernel are analyzed. A chirp-parameter design criterion is then derived to balance inverse-transform stability and multipath separation capability. Simulation results show that the proposed scheme achieves better bit-error-rate performance and stronger multipath robustness than conventional optical OFDM schemes in low-dynamic multipath conditions.

  (2)For high-dynamic underwater scenarios affected by Doppler spread due to relative motion and phase noise in coherent architectures, a coherent AFDM with index modulation (AFDM-IM) scheme is proposed. A system model incorporating multipath, Doppler shifts, and Wiener phase noise is established, and the equivalent input--output relationship in the discrete affine Fourier transform domain is derived. By exploiting the silent structure formed by inactive AFDM-IM positions as an implicit reference, a blind-search method driven by leakage-energy minimization and sparsity metrics is developed to accurately compensate the dominant residual carrier frequency offset without any additional pilot overhead, while maintaining robust tolerance to Wiener phase noise. Simulation results show that the proposed scheme outperforms the benchmark OFDM-IM scheme in terms of bit error rate under high-dynamic conditions and exhibits good phase-noise robustness within a practical range of laser linewidths.

  In summary, this thesis studies waveform construction, parameter design, and detection methods of AFDM for UWOC from the perspectives of low-dynamic and high-dynamic representative scenarios. The results demonstrate the feasibility of scenario-oriented AFDM adaptation under different mobility conditions and provide theoretical support for the design of reliable underwater optical communication systems.
\end{abstractEn}
